---
title: "Dimensionality Reduction"
author: "Mark Ngendo"
date: "7/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Relevance

The link to the dataset can be found here : http://bit.ly/CarreFourDataset

The column description is as follows:

* Invoice id: Computer generated sales slip invoice identification number

* Branch: Branch of supercenter (3 branches are available identified by A, B and C).

* City: Location of supercenters

* Customer type: Type of customers, recorded by Members for customers using member card and Normal for without member card.

* Gender: Gender type of customer

* Product line: General item categorization groups - Electronic accessories, Fashion accessories, Food and beverages, Health and beauty, Home and lifestyle, Sports and travel

* Unit price: Price of each product in $

* Quantity: Number of products purchased by customer

* Tax: 5% tax fee for customer buying

* Total: Total price including tax

* Date: Date of purchase (Record available from January 2019 to March 2019)

* Time: Purchase time (10am to 9pm)

* Payment: Payment used by customer for purchase (3 methods are available – Cash, Credit card and Ewallet)

* COGS: Cost of goods sold

* Gross margin percentage: Gross margin percentage

* Gross income: Gross income on a scale of 1000

* Rating: Customer stratification rating on their overall shopping experience (On a scale of 1 to 10)

## Libraries Needed
```{r}
library(ggplot2)
library(Rtsne)
library(FSelector)
library(e1071)
library(CatEncoders)
library(lattice)
library(corrplot)
library(caret)
```

## Data Reading

```{r loading data}

df <- read.csv("http://bit.ly/CarreFourDataset")
head(df)
```
### Structure of the data
```{r}
str(df)
```
## Data Cleaning
```{r}
# checking for duplicates
df[duplicated(df), ]
```
```{r}
# checking for missing values

colSums(is.na(df))
```



```{r}
# Accuracy

df$Invoice.ID <- NULL
df$Branch <- as.factor(df$Branch)
df$Customer.type <- as.factor(df$Customer.type)
df$Gender <- as.factor(df$Gender)
df$Product.line <- as.factor(df$Product.line)
df$Payment <- as.factor(df$Payment)
df$Date <- as.Date(df$Date, format = "%m/%d/%y")

```

## Exploaratory Data Analysis

### Univariate Analysis

```{r}
# creating a mode function
mode <- function(x){
  uniqx <- unique(x)
  uniqx[which.max(tabulate(match(x, uniqx)))]
}
```

#### Branch

```{r}

# Visualization
ggplot(df, aes(Branch)) +
  geom_bar(stat="count") +
  labs(title="Branches Distribution") +
  theme_classic()
```

Branch distribution is roughly equal

#### Gender
```{r}
# Visualization
ggplot(df, aes(Gender)) +
  geom_bar(stat="count")
```


The gender distribution in the dataset is balanced.

#### Customer type
```{r}
# visualization
ggplot(df, aes(Customer.type)) +
  geom_bar()
```

The customer types were balanced in terms of distribution.

#### Product Line
```{r}
# visualization
ggplot(df, aes(Product.line)) +
  geom_bar()
```
 Fashion Accessories and, Food and Beverage tie for the most bought categories but the distribution does
 not suggest an imbalance in general.
 
#### Unit Price
```{r}
# Mean
uprice.mean <- mean(df$Unit.price)
uprice.mean

# Mode
uprice.mode <- mode(df$Unit.price)
uprice.mode

# Median
uprice.median <- median(df$Unit.price)
uprice.median

# Standard Deviation
uprice.sd <- sd(df$Unit.price)
uprice.sd

# Kurtosis
uprice.kurt <- kurtosis(df$Unit.price)
uprice.kurt

# SKewness
uprice.skew <- skewness(df$Unit.price)
uprice.skew

# Range
uprice.range <- range(df$Unit.price)
uprice.range

# Visualization
ggplot(df, aes(Unit.price)) +
  geom_boxplot(outlier.colour = "red")
```


#### Quantity
```{r}
# mean
quantity.mean <- mean(df$Quantity)
quantity.mean

# Mode
quantity.mode <- mode(df$Quantity)
quantity.mode

# Median
quantity.median <- median(df$Quantity)
quantity.median

# Standard Deviation
quantity.sd <- sd(df$Quantity)
quantity.sd

# Range
quantity.range <- range(df$Quantity)
quantity.range

# Kurtosis
quantity.kurt <- kurtosis(df$Quantity)
quantity.kurt

# Skewness
quantity.skew <- skewness(df$Quantity)
quantity.skew

# Quantiles
quantity.quants <- quantile(df$Quantity)
quantity.quants

# Visualization
ggplot(df, aes(Quantity)) +
  geom_boxplot(outlier.colour = "red")
```


#### Tax
```{r}
# mean
tax.mean <- mean(df$Tax)
tax.mean

# mode
tax.mode <- mode(df$Tax)
tax.mode

# Median
tax.median <- median(df$Tax)
tax.median

# Standard Deviation
tax.sd <- sd(df$Tax)
tax.sd

# Kurtosis
tax.kurt <- kurtosis(df$Tax)
tax.kurt

# Skewness
tax.skew <- skewness(df$Tax)
tax.skew

# Range
tax.range <- range(df$Tax)
tax.range

# Quantiles
tax.quantiles <- quantile(df$Tax)
tax.quantiles

# Visual
ggplot(df, aes(Tax)) +
  geom_boxplot(outlier.colour = "red")
```

#### Date
```{r}
# mode
date.mode <- mode(df$Date)
date.mode

# median
date.median <- median(df$Date)
date.median

# standard deviation
date.sd <- sd(df$Date)
date.sd

```

#### Payment
```{r}

# Mode
payment.mode <- mode(df$Payment)
payment.mode



# visual
ggplot(df, aes(Payment)) +
  geom_bar(stat="count")
```

There is a fair distribution in the payment variable. However, fewer people tend to pay by Credit Card in these stores

#### COGS
```{r}
# mean
cogs.mean <- mean(df$cogs)
cogs.mean

# mode
cogs.mode <- mode(df$cogs)
cogs.mode

# median
cogs.median <- median(df$cogs)
cogs.median

# standard deviation
cogs.sd <- sd(df$cogs)
cogs.sd

# range
cogs.range <- range(df$cogs)
cogs.range

# kurtosis
cogs.kurt <- kurtosis(df$cogs)
cogs.kurt

# skewness
cogs.skew <- skewness(df$cogs)
cogs.skew

# quantiles
cogs.quantiles <- quantile(df$cogs)
cogs.quantiles

# visual

ggplot(df, aes(cogs)) +
  geom_boxplot(outlier.colour = "red")

```



#### Gross Income
```{r}
gi <- df$gross.income

# mean
gi.mean <- mean(gi)
gi.mean

# mode
gi.mode <- mode(gi)
gi.mode

# median
gi.median <- median(gi)
gi.median

# range 
gi.range <- range(gi)
gi.range

# standard deviation
gi.sd <- sd(gi)
gi.sd

# kurtosis
gi.kurt <- kurtosis(gi)
gi.kurt

# skewness
gi.skew <- skewness(gi)
gi.skew

# visual
 ggplot(df, aes(gross.income)) +
   geom_boxplot(outlier.colour = "red")

```



#### Rating
```{r}
# mean
rate.mean <- mean(df$Rating)
rate.mean

# mode
rate.mode <- mode(df$Rating)
rate.mode

# median
rate.median <- median(df$Rating)
rate.median

# standard deviation
rate.sd <- sd(df$Rating)
rate.sd

# range
rate.range <- range(df$Rating)
rate.range

# quantiles
rate.quantiles <- quantile(df$Rating)
rate.quantiles

# kurtosis
rate.kurt <- kurtosis(df$Rating)
rate.kurt

# skewness
rate.skew <- skewness(df$Rating)
rate.skew

# visual
ggplot(df, aes(Rating)) +
  geom_boxplot(outlier.colour = "red")
```

#### Total
```{r}
# mean
total.mean <- mean(df$Total)
total.mean

# median
total.median <- median(df$Total)
total.median

# mode
total.mode <- mode(df$Total)
total.mode

# standard deviation
total.sd <- sd(df$Total)
total.sd

# range
total.range <- range(df$Total)
total.range

# kurtosis
total.kurt <- kurtosis(df$Total)
total.kurt

# skewness
total.skew <- skewness(df$Total)
total.skew

# quantiles
total.quantiles <- quantile(df$Total)
total.quantiles

# visual
ggplot(df, aes(Total)) +
  geom_boxplot(outlier.colour = "red" )
```


### Bivariate Analysis

```{r product line vs Total Price}
ggplot(df, aes(x=Product.line, y=Total)) +
  geom_point()
```

Fashion Accessories have the highest Total prices while health and beauty products have a relatively lower price.

```{r Gendered Expenses}
ggplot(df ,aes(Gender, Total)) +
  geom_point()
```

Total Price is equally distributed in terms of gender

```{r Payment vs Total Price}
ggplot(df, aes(Payment, Total)) +
  geom_point()
```

The payment methods are nearly identical for the total prices of items at checkouts with some more expensive ones being attributed with Credit card payments.

```{r Gross Income vs Total}
ggplot(df, aes(gross.income, Total)) +
  geom_point()
```

As expected, there is a perfect positive linear relationship with how much the total is at checkout with the consumers gross income. 

```{r Customer type vs Total}
ggplot(df, aes(Customer.type , Total)) +
  geom_point()
```

Members and non members have a nearly equal distribution in expenditure with Members having no visible breaks in prices.

```{r Tax vs Total}
ggplot(df, aes(Tax, Total)) +
  geom_point()
```

There is a direct linear relationship between tax and total price. As expected, the higher the tax on items, the more they cost.

```{r Unit Price vs Total}
ggplot(df, aes(Unit.price, Total)) +
  geom_point()
```

There are several positive linear relationships with the Unit Price variable: the higher it is the higher the total price is. More data would be needed to explain the different lines considering they represent outside factors that influence the relationship. A good example would be the type of products being of different types.





```{r correlation table}

cor(df[,unlist(lapply(df, is.numeric))])
```
The following variables show strong correlation:

* Unit Price to Tax and COGS
* Quantity to Tax and COGS
* Gross Income to Unit Price, Quantity, Tax, COGS, Total


```{r copy of the dataset}
copy <- df[, -c(8, 9, 12, 15)]
label <- df[, 15]
```


## APPLICATION OF t-SNE ALGORITHM

```{r Label Encoding categorical columns}
branch <- LabelEncoder.fit(copy$Branch)
copy$Branch <- transform(branch, factor(df$Branch))

gender <- LabelEncoder.fit(copy$Gender)
copy$Gender <- transform(gender, factor(df$Gender))

customer <- LabelEncoder.fit(copy$Customer.type)
copy$Customer.type <- transform(customer, factor(copy$Customer.type))

product <- LabelEncoder.fit(copy$Product.line)
copy$Product.line <- transform(product, factor(copy$Product.line))

pay <- LabelEncoder.fit(copy$Payment)
copy$Payment <- transform(pay, factor(copy$Payment))

```

```{r building the model}
model <- Rtsne(copy, dims=2, perplexity=30, verbose= TRUE, max_iter=1000)
summary(model)
```
```{r viewing output}
head(model$Y)
```

```{r visualizing results}

plot(model$Y, t='p', main="Output of TSNE")
```


## Feature Selection

```{r correlation method}
corrMat <- cor(copy)

# highly correlated features
high <- findCorrelation(corrMat, cutoff = .75)

# names of highly correlated features

names(copy[, high])
```


```{r comparison after dropping high}
# data set without highly correlated variables
c2 <- copy[-high]

par(mfrow = c(1, 2))

# plotting
corrplot(corrMat, order = "hclust")
corrplot(cor(c2), order = "hclust")
```



```{r Feature Ranking method}
# adding Ratings column to copy data set

copy$Rating <- df$Rating

# how different variable score in relation to rating variable

scores <- information.gain(Rating~., copy)
subset <- cutoff.k(scores, 5)

# Needed features (5)
as.data.frame(subset)
```

## Conclusions


From the above analysis,  the following variables have been found to be highly correlated:

* Tax
* Cost of Goods
* Gross Income


Using a Feature Ranking method with Information Gain of all variables being used as a metric of comparison, 
the following five variables would be the best to use for modeling a regressor with respect to Rating:

* Branch
* Customer Type
* Gender
* Product Line
* Unit Price

## Recommendations
The following are my recommendations based on the above analysis:

* The marketing department should focus on the Electronics, Health and Beauty and Lifestyle departments in terms of promotions seeing as the data says that these areas are the smaller contributors.

* To offer some added benefits that come with being a member seeing as the analysis shows that members are distributed at all price points whereas non members have some minute gap. This would help to motivate more people to become members and hopefully increase the sales.

To use the identified features in the conclusions segment if they want to build a regression model to predict the sales. These features have been curated extensively and can yield better results than the entire roster of variables provided. 